---
title: " "
listing: _posts
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)

# Attaching relevant packages:

library(tidyverse) # Wickham et al. 2019
library(here) # Müller 2020
library(sf) # Pebesma 2018
library(raster) # Hijmans 2020
library(fasterize) # Ross 2020
library(rgdal) # Bivand et al. 2021
library(rnaturalearth) # South 2017a
library(rnaturalearthdata) # South 2017b
```

# Data Project Example 3: Working With Raster Data

**Project Description**

*In this simple data project, AquaMaps raster layers detailing the probability of occurrence (richness) - based on relative environmental suitability, including species preferences for water temperature, depth, salinity, and distance to land - of 35 cetacean species that can be found in the Southern California Bight, are explored and visualized via a single static map. The extent of the raster map is 122°W to 116°W and 31°N to 35°N. To access an .html file with project code, click [here](working_raster_data.html).*

```{r}
# Reading in a CA coastline shapefile from the 'rnaturalearth' package:

ca_coastline <- ne_coastline(scale = "medium", 
                             returnclass = "sf")

# Reading in the raster layers all at once:

ca_cetaceans <- list.files(path = here("ca_cetaceans"), 
                    pattern = ".tif", 
                    full.names = TRUE)

# Stacking all layers as a multilayer:

cetaceans_stacked <- stack(ca_cetaceans)

# Defining the algebra on multiple raster layers:

cetaceans_prob <- function(x) {
  if_else(x > 0.9, 
          1, 
          0)
}

cetaceans_occurrence <- calc(cetaceans_stacked, 
                             function(x) {
                               cetaceans_prob(x)
                               })

cetaceans_summed <- calc(cetaceans_occurrence, 
                         sum)

# Converting the rasters to a data frame:

cetaceans_df <- rasterToPoints(cetaceans_summed) %>% 
  as.data.frame()

# Graphing the raster data frame:

ggplot(data = ca_coastline) + 
  geom_sf() + 
  coord_sf(xlim = c(-122, 
                    -116), 
           ylim = c(31, 
                    35), 
           expand = FALSE) + 
  geom_raster(data = cetaceans_df, 
              aes(x = x, 
                  y = y, 
                  fill = layer), 
              alpha = 0.8) + 
  theme_void() + 
  scale_fill_gradientn(colors = c("darkseagreen1",
                                  "seagreen3", 
                                  "seagreen"), 
                       name = "Cetacean spp.\nRichness") + 
  labs(title = "Cetacean Species Richness Off the Coast of the Southern CA Bight") + 
  theme(plot.title = element_text(face = "bold", 
                                  vjust = 2.5))
```

**Figure 3.1.** A raster showing species (spp.) richness - the number of species that could be found in a particular location - of 35 cetaceans off the coast of the Southern California Bight. The color gradient calculates richness based on the relative environmental suitability and species preferences for water temperature, depth, salinity, and distance to land. The numerical threshold used was 0.9, to provide a greater gradient extent of species richness *(Source: Kaschner et al. 2016)*.

**Citations**

Bivand, R., Keitt, T., & Rowlingson, B. (2021). rgdal: Bindings for the 'Geospatial' Data Abstraction Library [R package].

Hijmans, R.J. (2020). raster: Geographic data analysis and modeling [R package].

Kaschner *et al.* (2016). *AquaMaps: Predicted range maps for aquatic species* [Raster data].

Müller, K. (2020). here: A simpler way to find your files [R package].

Pebesma, E. (2018). Simple features for R: Standardized support for spatial vector data [R package]. *The R Journal*, *10*(1), 439-446.

Ross, N. (2020). fasterize: Fast polygon to raster conversion [R package].

RStudio Team. (2021). *RStudio: Integrated development environment for R* (Version 4.0.3) [Computer software]. RStudio PBC, Boston, USA.

South, A. (2017a). rnaturalearth: World map data from Natural Earth [R package].

South, A. (2017b). rnaturalearthdata: World vector map data from Natural Earth used in 'rnaturalearth' [R package].

Wickham *et al.* (2019). Welcome to the tidyverse [R package]. *Journal of Open Source Software*, *4*(43), 1686.

^*webpage contains ip @ 2021 Carlos Simms*^