---
title: " "
listing: _posts
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE)

# Attaching relevant packages:

library(tidyverse) # Wickham et al. 2019
library(here) # M체ller 2020
library(janitor) # Firke 2021
library(sf) # Pebesma 2018
library(tmap) # Tennekes 2018
library(gstat) # Pebesma 2004, Gr채ler et al. 2016
library(stars) # Pebesma 2021
```



# Data Project Example 2: Visualizing Spatial Data



**I. Locations of 2008 Oil Spill Events in CA**

```{r}
# Reading in the "CA DFW Oil Spill Incident Tracking" data:

ca_oil_spills <- read_sf(here("ca_oil_spills"), 
                         layer = "ds394") %>% 
  clean_names()

# Creating an exploratory interactive 'tmap':

tmap_mode("view")

tm_shape(ca_oil_spills) + 
  tm_dots("red")
```

**Figure 2.1.** An interactive map of California detailing the exact locations of all recorded 2008 oil spill events in and off the coast of the state *(Source: Cal EMA & DFG-OSPR 2009)*.

**II.**

```{r}
# Reading in the CA county shapefile data:

ca_counties <- read_sf(here("ca_counties"), 
                       layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name) %>% 
  rename(county_name = name)

# Wrangling the data to specify the total count of inland oil spill events by county:

ca_inland_spills <- ca_oil_spills %>% 
  filter(inlandmari == "Inland") %>% 
  count(localecoun) %>% 
  rename(count = n)

# Matching the CRS of the two data frames and then joining:

ca_inland_spills <- st_transform(ca_inland_spills, 
                                 3857)

ca_spill_counts <- ca_counties %>% 
  st_join(ca_inland_spills)
  
# Creating a finalized static chloropleth map:

ggplot(data = ca_spill_counts) +
  geom_sf(aes(fill = count), 
          color = "darkred", 
          size = 0.1) +
  theme_void() + 
  scale_fill_gradientn(colors = c("yellow", 
                                  "red", 
                                  "darkred"), 
                       name = "Total Number\nof Oil Spills") + 
  labs(title = "2008 Oil Spill Events in CA Occurring Inland by County") + 
  theme(plot.title = element_text(face = "bold", 
                                  hjust = 0.5, 
                                  vjust = 2.5))
```

**Figure 2.2.** A static chloropleth map of California detailing the total counts of all recorded 2008 oil spill events in the state, occurring inland by county *(Source: Cal EMA & DFG-OSPR 2009)*.

**Citations**

Cal EMA & DFG-OSPR. (2009). *Oil spill incident tracking* (Edition 2008) [Raw data]. California Department of Fish and Game, Office of Spill Prevention and Response, Sacramento, USA.

Firke, S. (2021). janitor: Simple tools for examining and cleaning dirty data [R package].

Gr채ler, B., Pebesma, E., & Heuvelink, G. (2016). Spatio-temporal interpolation using gstat [R package]. *The R Journal*, *8*(1), 204-218.

M체ller, K. (2020). here: A simpler way to find your files [R package].

Pebesma, E. (2004). Multivariable geostatistics in S: The gstat package [R package]. *Computers & Geosciences*, *30*, 683-691.

Pebesma, E. (2018). Simple features for R: Standardized support for spatial vector data [R package]. *The R Journal*, *10*(1), 439-446.

Pebesma, E. (2021). stars: Spatiotemporal arrays, raster and vector data cubes [R package].

RStudio Team. (2021). *RStudio: Integrated development environment for R* (Version 4.0.3) [Computer software]. RStudio PBC, Boston, USA.

Tennekes, M. (2018). tmap: Thematic maps in R [R package]. *Journal of Statistical Software*, *84*(6), 1-39.

Wickham *et al.* (2019). Welcome to the tidyverse [R package]. *Journal of Open Source Software*, *4*(43), 1686.
